
---
title: "Causal Impact - campaign analysis"

output: rmarkdown::github_document
    toc: true
---

# 

#### This analysis uses 2 data sets
##### 1. Daily transactions: 
###### Aim: To detect trend, seasonality in the time series data

Query:
      
      SELECT o.parent_vertical,
            p.created_at,
            sum(no_transactions) no_transactions

      from `gc-prd-ext-data-test-prod-906c.gc_paysvc_live.organisations` o 
      inner join `gc-prd-ext-data-test-prod-906c.gc_paysvc_live.mandates` m
      on o.id = m.organisation_id
      left join 
                (select mandate_id, 
                        cast(created_at as date) as created_at,
                        count(id) no_transactions 
                 from `gc-prd-ext-data-test-prod-906c.gc_paysvc_live.payments` p
                 where 
                     amount > 0  and created_at is not null and charge_date is not null --successfully transaction
                 group by mandate_id,currency,cast(created_at as date)  ) p
      on p.mandate_id = m.id

      where p.created_at is not null

      group by o.parent_vertical,  p.created_at


##### 2. Cummulative number of activations: 
###### Aim: To compare pre and post campaign 


Query:

    select 
        distinct t1.created_at, 
        sum(count(t1.mandate_id)) over ( order by t1.created_at asc) as cum_activated_mandates,
        case when t1.created_at < cast('2016-12-01' as date) 
        then 0 else 1
        end as campaign_status  ---0 pre, 1 post
    from 
    (
              select 
                   min(cast(p.created_at as date)) as created_at,   
                   mandate_id

              from `gc-prd-ext-data-test-prod-906c.gc_paysvc_live.payments` p
              inner join  `gc-prd-ext-data-test-prod-906c.gc_paysvc_live.mandates` m
              on m.id = p.mandate_id

              where p.created_at is not null
              and p.amount > 0 and p.created_at is not null and p.charge_date is not null --successfully transaction
              group by mandate_id
            ) 
            as t1

    group by t1.created_at

```{r}
install.packages(c("CausalImpact",
                   "ggfortify",
                    "changepoint",
                    "strucchange",
                    "ggpmisc",
                    "ggeasy")
                )
```


```{r}
library(CausalImpact)
library(tidyverse)
library(xts)
library(zoo)
library(lubridate)
library(SparkR)
library(tidyr)
library(dplyr)
library(ggeasy)
require(gridExtra)
```


```{r}
# Cummulative activation rate data set
act <- read.df("dbfs:/FileStore/shared_uploads/phuong.le@man-es.com/CausualImpact_all.csv",
                "csv",
                header=TRUE)

# Daily transaction data set
daily <- read.df("dbfs:/FileStore/shared_uploads/phuong.le@man-es.com/dailytransaction.csv", 
                 "csv", 
                 header=TRUE)
act <- collect(act)
daily <- collect(daily)
```


```{r}
# Explore data
daily$no_transactions <-as.numeric(daily$no_transactions)
daily$created_at<-as.Date(daily$created_at)
str(daily)
```


```{r}
transaction_plot <- ggplot(data = daily, aes(created_at, 
                         no_transactions)) +
                          geom_line(color = "lightblue", size = 0.5) +
                          geom_point(color = "red", size = 0.05) + 
                          labs(title = "Number of transactions on daily basis",
                          y = "Transactions", x = "") + 
                          facet_grid( parent_vertical ~.,scales="free") 
transaction_plot 
```

# 

##### Comments about the plot above:

######High growth in number of transactions pre vs post campaign can be seen in:
    Health care
    Sport fitness
    Tradesmen
    Digital service
######Low growth or no growth at all, indicating campaign seems not to influence much in: 
    Prefessional and financial service
    Societies and club
    Property
    
######Except sociaties and club, all other verticals have seasonal impact over time.
######Seasonality can be seen in 
1. Monthly season
    Professional and financial services 
    Sport fitness
2. Combined monthly and weekly season
    Property
    Digital services
    Tradesmen and non professionals services
3. Weekly season
    Healthcare

=> These observations will be used to highlight seasonality in the times series model in later part. 

##### Causal Impact model

This implementation use causal model using Bayesian structural time-series model

Use 2 libraries bsts and CasualImpact to measure the causal effect of the campaign on a time series.

Source:

  https://cran.r-project.org/web/packages/CausalImpact

```{r}
# Explore cummulative activation rate dataset
vars <-c("created_at","cum_activated_mandates","campaign_status")
pre.period <- c(as.Date("2016-04-02"), as.Date("2016-11-30"))
post.period <- c(as.Date("2016-12-01"),as.Date("2017-03-06"))
```


```{r}
act$created_at <- as.Date(act$created_at)
act$cum_activated_mandates <- as.numeric(act$cum_activated_mandates)
act$campaign_status  <- as.numeric(act$campaign_status)
str(act)
```


```{r}

cum_activated_mandates<-xts(as.numeric(act$cum_activated_mandates),
                            as.Date(act$created_at))

campaign_status<-xts(as.numeric(act$campaign_status),
                     as.Date(act$created_at))

act_ts <- cbind(cum_activated_mandates,campaign_status)
act_ts <-zoo(act_ts)

head(act_ts)
```


```{r}
# Fit with month seasonal
impact4 <- CausalImpact(act_ts, 
                          pre.period, 
                          post.period,
                          model.args = list(niter = 1000,
                                            nseasons = 365, 
                                            season.duration =7))
plot(impact4)
```


```{r}
summary(impact3)
```

